---
title: "Annotating"
author: "Mikias HW"
date: "`r Sys.Date()`"
output: html_document
---
# Prep Workspace
## Load libraries
```{r message=FALSE, warning=FALSE}
library(Seurat)
library(SeuratData)
library(patchwork)
library(conflicted)  # Resolve function name conflicts
library(tidyverse)  # Data manipulation and visualization
library(cowplot)
theme_set(theme_cowplot())
```

## Conflicted functions
```{r results='hide', message=FALSE, warning=FALSE}
# Set function preferences
conflicts_prefer(dplyr::select)

conflicts_prefer(dplyr::filter)
```

## Define Directories
```{r}
# Define data directory
dataDir <- paste(getwd(), "data", sep = "/")

list.files(dataDir)

# Define output directory
outDir <- paste(getwd(), "output", "annotating", sep = "/")
```

# Load Data
```{r}
panc8 <- LoadData("panc8")
table(panc8$tech)
```

# Quick Processing
```{r}
# we will use data from 2 technologies for the reference
pancreas.ref <- subset(panc8, tech %in% c("celseq2", "smartseq2"))
pancreas.ref[["RNA"]] <- split(pancreas.ref[["RNA"]], f = pancreas.ref$tech)

# pre-process dataset (without integration)
pancreas.ref <- pancreas.ref |> 
  NormalizeData() |> 
  FindVariableFeatures() |> 
  ScaleData() |> 
  RunPCA() |> 
  FindNeighbors(dims = 1:30) |> 
  FindClusters() |> 
  RunUMAP(dims = 1:30)
```

# View
```{r}
DimPlot(pancreas.ref, group.by = c("celltytpe", "tech"))
```

# Integrate
```{r}
pancreas.ref <- IntegrateLayers(object = pancreas.ref, 
                                method = CCAIntegration, 
                                orig.reduction = "pca",
                                new.reduction = "integrated.cca", 
                                verbose = FALSE)

pancreas.ref <- pancreas.ref |> 
  FindNeighbors(reduction = "integrated.cca", dims = 1:30) |> 
  FindClusters() |> 
  RunUMAP(reduction = "integrated.cca", dims = 1:30)

```

# View
```{r}
DimPlot(pancreas.ref, group.by = c("tech", "celltype"))
```

# Classificaiton with integrated data
```{r}
# select two technologies for the query datasets
pancreas.query <- subset(panc8, tech %in% c("fluidigmc1", "celseq"))

pancreas.query <- NormalizeData(pancreas.query)

pancreas.anchors <- FindTransferAnchors(reference = pancreas.ref, 
                                        query = pancreas.query, dims = 1:30,
                                        reference.reduction = "pca")

predictions <- TransferData(anchorset = pancreas.anchors, 
                            refdata = pancreas.ref$celltype, dims = 1:30)

pancreas.query <- AddMetaData(pancreas.query, metadata = predictions)
```

# Verify matching
```{r}
pancreas.query$prediction.match <- pancreas.query$predicted.id == pancreas.query$celltype

table(pancreas.query$prediction.match)

table(pancreas.query$predicted.id)
```

# View
```{r}
VlnPlot(pancreas.query, c("REG1A", "PPY", "SST", "GHRL", "VWF", "SOX10"), group.by = "predicted.id")
```

# Unimodal UMAP Projection
```{r}
pancreas.ref <- RunUMAP(pancreas.ref, dims = 1:30, reduction = "integrated.cca", return.model = TRUE)

pancreas.query <- MapQuery(anchorset = pancreas.anchors, 
                           reference = pancreas.ref, 
                           query = pancreas.query,
                           refdata = list(celltype = "celltype"), 
                           reference.reduction = "pca", 
                           reduction.model = "umap")
```
MapQuery() is a wrapper around three functions: TransferData(), IntegrateEmbeddings(), and ProjectUMAP(). TransferData() is used to transfer cell type labels and impute the ADT values; IntegrateEmbeddings() is used to integrate reference with query by correcting the query’s projected low-dimensional embeddings; and finally ProjectUMAP() is used to project the query data onto the UMAP structure of the reference. The equivalent code for doing this with the intermediate functions is below:
```{r}
pancreas.query <- TransferData(anchorset = pancreas.anchors, 
                               reference = pancreas.ref, 
                               query = pancreas.query,
                               refdata = list(celltype = "celltype"))

pancreas.query <- IntegrateEmbeddings(anchorset = pancreas.anchors, 
                                      reference = pancreas.ref, 
                                      query = pancreas.query,
                                      new.reduction.name = "ref.pca")

pancreas.query <- ProjectUMAP(query = pancreas.query, 
                              query.reduction = "ref.pca", 
                              reference = pancreas.ref,
                              reference.reduction = "pca", 
                              reduction.model = "umap")
```

# View both datasets
```{r}
p1 <- DimPlot(pancreas.ref, reduction = "umap", group.by = "celltype", label = TRUE, label.size = 3,
    repel = TRUE) + NoLegend() + ggtitle("Reference annotations")

p2 <- DimPlot(pancreas.query, reduction = "ref.umap", group.by = "predicted.celltype", label = TRUE,
    label.size = 3, repel = TRUE) + NoLegend() + ggtitle("Query transferred labels")

p1 + p2
```

```{r}
sessionInfo()
```

